Optimal recommender system with strategic alternatives \\ Market Design. Term Project

### 2019-12-24 Draft.tex
Draft

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=15mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\begin{document} %

	\section{Question}
	Suppose we have a web-platform that is designed to implicitly match customers and service providers via the past-user-based recommendation (examples of such a kind of platforms include \textit{TripAdvisor, Yelp, etc}). The main goal of the platform is to render users satisfied with the quality of consumed products (in order to keep users using the platform). Users' reviews help the platform to infer about the qualities of different providers and based on this, to recommend users better options. However, once a new provider appears, the platform usually does not know the quality of the good provided, since there are no reviews yet. And the natural desire of the platform in this case is to figure out the quality of new alternative as quickly as possible in order to use (or not) it in recommendation for users. But how can the platform achieve this? Well, it seems that the platform can just start recommend the new alternative to users, collect reviews and then infer about the quality. But obviously the platform cannot do it too often and with too risky options : users will just stop using the platform that recommends them things that eventually turn out to be inferior. And here the first trade-off arises: how to induce exploration of new alternative without "sacrifice" customers too much? There are several papers on this topic that shed light on this trade-off and propose the optimal policy. The main idea in all those papers is that there is an optimal level of information obfuscation to create incentives for exploring new things.

	However, in all previous works the quality of alternatives was exogenous to the process of recommendation. But in real life, providers are also strategic agents (and in most cases they are way more sophisticated than customers) who can change quality of their products in response to the recommendation policy the platform implements, and taking this into account the recommendation policy itself (I think) should change! So, questions I would be interested to address are: \textbf{what would be the optimal information revelation (recommendation) policy in this case, and how it is affected by the presence of strategic providers? Whether the quality of products increases (compare to the case when alternatives are exogenous)?}

	With strategic alternatives the platform's and customers' preferences may become even more misaligned. Because now customers may be used by the platform not only for the sake of exploration (and to the benefit of subsequent agents) but also as a tool for encouraging the competition between alternatives.

	\section{Model}
	\subsection{In words}
	We have a platform that is designed for providing customers with some services, and two firms (providers) are operating on this platform (it can be restaurants, hotels etc.).

	 Customers arrive sequentially and each should decide on the product from which firm to choose. Although the qualities of the products are unknown \textit{ex ante}, customers have some priors and make their decision rationally taking into account all information available and updating their beliefs via the Bayes rule. After making a decision, each customer gets some reward, reports his experience to the platform (principal), and leaves the system.

	The platform accumulates agents' reviews and based on this can recommend a product from a particular firm to the subsequent agents. The goal of the platform is to maximize the (infinite horizon ?) expected discounted reward of all agents.

	Heterogeneous firms, in turn, are competing for agents and can invest to improve quality of their goods. The goal of each firm is to maximize its (infinite horizon ?) expected discounted profit taking into account the recommendation policy.

	At $t = 0$ each firm privately learns its type. Then, the timing of the $t$-th ($t=1, 2, \dots$) stage of the game is the following:
	\begin{enumerate}[1)]
		\item Each firm decides whether to stay in business or not (and incurs cost $f_i$ if decides to stay)
		\item Firms make private and costly investment in quality
		\item The unobserved quality is realized
		\item Customer arrives
		\item The principal makes a recommendation
		\item Customer decides which firm to choose
		\item Payoffs are realized and the principal observes customers' experience
	\end{enumerate}
	\subsection{Formalization}
	\subsubsection{Firms}
	We have two competing firms. Each firm has its own type that somehow is supposed to represent its "effectiveness" or "inclination to investing and sustaining the high quality of the good". The simplest example that comes in mind is that the type represents the firm's marginal cost of investment: the higher the type - the less effective the firm is. (types are either discrete $\theta^H, \theta^L$ or continuous $\theta \in (0, 1)$).

	\textbf{Question:} why do we need types?

	\textbf{Answer:} customers will learn and infer about the type of the firm

	Making investment $e \in \mathbb{R}_{+}$ (maybe make it discrete $e^L$ and $e^H$) in quality of its product, each firm incurs cost $c_i(e)$ that should depend on the firm's type (for example $c_i(e) = \theta_ie_i$ where $\theta_i$ is the type of the $i$-th firm). Moreover, if the firm decides to stay in business, it incurs fixed cost $f_i$ disregard whether the customer chooses the firm or not. $f_i$ is low enough to ensure that at least one firm will operate in any case (maybe fixed cost also should dependent on type).
	Each firm obtains a profit $1$ if the customer chooses its product. Denote for each firm $$d_{it} = \begin{cases}
	1, \text{if consumer }t\text{ chooses the firm }i\\
	0, \text{if consumer }t\text{ does not choose the firm }i\\
	\end{cases}$$
	and
	$$b_{it} = \begin{cases}
	1, \text{if the firm }i\text{ decided to stay in busimess at }t\\
	0, \text{if the firm }i\text{ decided not to stay in busimess at }t
	\end{cases}$$
	then each firm is maximizing $$\underset{\{e_{it}, b_{it}\}_{t=1}^{\infty} }{\max}\ \expect \left[\sum_{t=1}^{\infty} \delta^t b_{it}(d_{it} - c_i(e_{it})- f_i)\right]$$
	subject to the recommendation policy implemented by the principal.

	\textbf{Open questions:}

	can the firms observe the experience of users who chose the competitor? It seems reasonable if they could : the restaurant manager can send somebody to the nearby competitive restaurant to infer about the real quality.

	can the firm know that the competitor still in business?

	\underline{Maybe infinite-horizon model will be too hard, make it finite (start from two-periods)}

	\subsubsection{Products}
	Each firm provides a product which in time $t$ has quality $q_{it} \in (0, 1)$  which is endogenously formed from the firms' investment $e_{it}$ in this period.

	\textbf{Idea:} the quality should somehow represent consumers' satisfaction of the product, the higher quality - the higher (expected) satisfaction.

	\textbf{Realization:} Suppose that upon consuming the good from the firm $i$, the customer $t$ gets a reward $$r_{it} = \begin{cases}
	1, \text{ w.p. } q_{it}\\
	0, \text{ w.p } 1 - q_{it}
	\end{cases}$$
    The quality of the product from the firm $i$ at time $t$ is determined as $$q_{it} = \begin{cases} \frac{e_{it}}{e_{it}+e_{jt}},\ &j \neq i, e_{it}^2 + e_{jt}^2 \neq 0 \\
	0, &e_{it} = e_{jt} = 0
	\end{cases}
	$$

	\textbf{Question:} why is the quality of the alternative $i$ affected (negatively) by the investment made by the firm $j$?

	\textbf{One possible answer:} in real life there is no such a thing as an absolute quality. I think that good things are not good \textit{per se}, they are good \underline{in comparison} to other things. So, in this sense the assumption may seem to be reasonable.

	\textbf{Concern:} in this set-up if both firms make high and equal investment the probability of success for the customer is $\frac{1}{2}$, whereas if only one firm made high investment while the other made low, the probability of success for the first firm becomes higher than $\frac{1}{2}$. Does it represent the reality?

	\subsubsection{Principal}
	Principal observes the history of consumers' choices and their payoffs, $h^t$ (denote $H^t$ the whole set of such histories) and can commit to the recommendation policy, a function $R: H^t \to M^t$, where $M^t$ is the message space feasible for the customer $t$. The principal's goal is to design a revelation policy $R(\cdot)$ in order to maximize agents' expected discounted payoff:
	$$\pi = \underset{R(\cdot)}{\max}\ \expect\left[\sum_{t=1}^{\infty} \delta^t r_{t} \right]$$

	where $r_t$ is an expected reward of the customer $t$.

	\subsubsection{Customers}

	Customers (agents) arrive sequentially one by one, they do not observe the experience of other agents, but can receive information revealed by the principal. For each principal's message $m_t \in M^t$ the agent's strategy is a function $\sigma^t: M^t \to A$, where $A = \{a_1, a_2\}$ describes the set of possible alternatives (products from the firm 1 or 2). In other words, the agent's strategy prescribes for each principal's message an action, the product from which firm to choose. Choosing the strategy, each agent $t$ maximizes
	$$\underset{\{a_1, a_2\}}{\max}\ \expect_t \left[q_{it}|m_t \right]$$

	\textbf{Questions}: what are priors? which alternative do the customers choose when indifferent between two?

\section{Analysis}
\subsection{Straightforward observations}
First of all, it seems that in this model it will never be optimal (starting from some point) to start recommending only one firm. This is because in this case the firm which is not recommended leaves the business (because of fixed cost) and this eliminates any incentive for the incumbent to invest. Hence, the quality becomes $0$ and all subsequent customers will get 0. So, the platform should internalize its role in sustaining a competition.
\subsection{No disclosure at all}
Suppose that the platfotm employs the non-disclosure policy. That is, agents do not possess any information regarding the previous users' experience. In this case, agents make decisions based on priors. If customers are homogeneous and have identical priors about the quality, then the firm which the customers are $a\ priori$ inclined towards stays in business with zero investment while the second firm exits the market. In this equilibrium all customers get zero, so the platform's payoff is zero.

As we can see, now with strategic agents the non-disclosure policy becomes much worse. One reason is that it completely eliminates any incentive to increase the quality of alternatives.

\subsection{Full disclosure}
\subsubsection{1 period, observable qualities}
Suppose we have only 1 period, at the beginning of the period, firms privately draw its types $\theta$ according to some full-support continuous distribution with the p.d.f. $f(\theta)$. Then both firms simultaneously perform investment. After that the customer arrives and can observe realized qualities. The customer chooses the firm with higher quality (ties do not happen almost surely).
In this case the following Lemma is true
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$ then
	in the one period simultaneous-move game with observable qualities there exist a symmetric equilibria in which both firms invest $e(\theta) = -\ln(\theta)$.
\end{lemma}
\begin{proof}
	See Appendix
\end{proof}

\subsubsection{Many periods, observable qualities}

If there are more than 1 period then in all periods after the first one the firm with higher type will learn that its type is high (because the customer did not choose it in the first period) and that is why the aforementioned equilibrium breaks down: less effective firm is better-off if it leaves the business, but if one of the firm leaves then the other has no incentive to invest: it becomes a monopoly. So let us note this result
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$ then in many-periods simultaneous-move game with observable qualities there exists an equilibrium in which:
	\begin{enumerate}
		\item In the first period both firms play $e(\theta) = -\ln(\theta)$
		\item In all subsequent periods:
		\begin{itemize}
			\item the more efficient firm (with smaller $\theta$) plays $e(\theta) = -\ln(\theta)$ if and only if there is a competitor in business, and $e(\theta) = 0$ otherwise
			\item the less efficient firm (with higher $\theta$) leaves the business forever.
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See Appendix.
\end{proof}
As we can see, if qualities are observable in many periods game, unravelling may happen, and the principal's expected payoff becomes equal to the expected reward of the first agent $$\pi_0 = r_1 = \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] = \frac{3}{4},\ i \neq j$$

\subsection{Not full disclosure}
\subsubsection{Qualities are observable by the principal}
Can the principal do better? Let us take the simplest case, when the principal can observe the qualities.
Suppose the principal commits to recommend high quality firm with probability $q$ and low probability good with probability $1-q$. In this case it is easy to see that the symmetric equilibrium of one-period game is an equilibrium in many-periods game
\begin{lemma}
	Suppose $\theta \sim U(0, 1)$, $f \le 0.5$, the platform can observe the realized qualities and commits to recommend to the customer better option with probability $q \in (0.5, 1)$ and worse option with probability $1-q$. Then there exists a symmetric equilibium in which both firms are playing $e(\theta) = -(2q - 1)\ln \theta$.

	Moreover, in this equilibrium the platform's expected payoff is strictly greater than the platform's expected payoff $\pi_0$ in case of unravelling (Lemma 2) 
\end{lemma}
\begin{proof}
	See Appendix.
\end{proof}

As we can see, the platform can encourage competition.

\section{Extensions}
\begin{enumerate}
	\item Agents do not know they place in line.
	\item Heterogeneous customers
\end{enumerate}
\subsection*{Appendix}

\begin{proof}[Proof of Lemma 1]
	Let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function. Then each firm is maximizing $$\underset{e}{\max}\ \prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) - \prob(\theta \le \varepsilon^{-1}(e))(\theta e + f)$$
	or using the fact that $$\prob(\theta \le x) = \begin{cases}
	1, x > 1\\
	x, x \in [0, 1]\\
	0, x < 0
	\end{cases}$$
	the firm is maximizing
	$$\underset{e}{\max}\ 1 - \varepsilon^{-1}(e) - \theta e - f$$
	F.O.C.:
	$$-\frac{1}{\varepsilon'(\varepsilon^{-1}(e))} = \theta$$
	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
	$$\varepsilon'(\theta) = -\frac{1}{\theta}$$
	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -\ln\theta$$ 
\end{proof}

\begin{proof}[Proof of Lemma 3]

	If both firms stay in business in the particular period, then let us find the equilibrium strategy for firms in a form $\varepsilon(\theta)$ where $\varepsilon(\theta):(0, 1) \to \mathbb{R}$ is a decreasing function. Then each firm is maximizing 
	\begin{align*}\underset{e}{\max}\ q\{\prob(\theta > \varepsilon^{-1}(e))(1 - \theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(-\theta e - f)\} +\\+ (1-q)\{\prob(\theta > \varepsilon^{-1}(e))(-\theta e - f) + \prob(\theta \le \varepsilon^{-1}(e))(1 - \theta e - f)\}
	\end{align*}
	or using the fact that $$\prob(\theta \le x) = \begin{cases}
	1, x > 1\\
	x, x \in [0, 1]\\
	0, x < 0
	\end{cases}$$
	the firm is maximizing 
	$$\underset{e}{\max}\ q(1 - \varepsilon^{-1}(e)) + (1 - q)\varepsilon^{-1}(e) - \theta e - f$$
	F.O.C.: $$-\frac{q}{\varepsilon'(\varepsilon^{-1}(e))} + \frac{1-q}{\varepsilon'(\varepsilon^{-1}(e))} - \theta = 0$$
	Since the equilibrium we are interested in is symmetric, $\varepsilon(\theta) = e$ and hence
	$$\varepsilon'(\theta) = -\frac{2q - 1}{\theta}$$
	Assuming $\varepsilon(1) = 0$ we get $$\varepsilon(\theta) = -(2q - 1)\ln\theta$$.

	It remains to check that less effective firm stays in business. Expected profit from staying in business for less effective firm is equal to
	$$\expect[1-q + (2q-1) \theta \ln \theta - f] = \frac{5}{4} - \frac{3}{2}q - f \ge 0$$

	The expected payoff of the platform is equal to
	$$\pi = \frac{q}{1-\delta} \expect\left[\frac{\ln \theta_i}{\ln \theta_i + \ln \theta_j}\bigg| \theta_i < \theta_j\right] +\frac{1-q}{1-\delta}\expect\left[\frac{\ln \theta_j}{\ln \theta_i + \ln \theta_j}\bigg| \theta_j \ge \theta_i\right] = \frac{2q+1}{4(1 - \delta)} > \frac{3}{4} = \pi_0$$
\end{proof}
\end{document}

>>>>

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=15mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\title{Optimal recommender system with strategic alternatives \\ Market Design. Term Project}
\date{}
\author{Danil Fedchenko}
\begin{document} % 
	\maketitle
	\section{Introduction}
	Recommender systems became an integral part of our lives. The most widely known examples of such a kind of services include: \textit{TripAdvisor} which presents ranking on Hotels and Restaurants for customers based on previous users' experience; \textit{Yelp} which provides a forum with reviews and ranking for different businesses; \textit{Netflix} which recommends movies and series also based on information aggregated from different customers. All those services are very important since, on one hand, they provide people with information about options they otherwise may never be aware of, at no (or very low) cost. On the other hand, these services aggregate the information and may make rather precise recommendations to people which significantly reduce search cost for them. But here the interesting question arise: how to disclose the information, the platform aggregated, optimally? It turns out that the full disclosure may not be optimal from the social welfare point of view. This happens since users, who do not internalize the information externalities they impose to other agents, underexplore i.e. start to choose the safe option and do not try new things.

	Several papers studied this question in different frameworks (see Literature review section) and the main conclusion that all those paper draw is that there should be an optimal level of information obfuscation in order to induce short-living and myopic users to experiment. However, in the previous work the quality of alternatives was supposed to be exogenous to the process of recommendation and learning. So, the recommender policy the platform implements cannot affect the realized quality of alternatives that users choose. But in reality, these alternatives are also strategic agents (think about restaurants or hotels), and in most cases these agents are way more sophisticated than users. These strategic agents may adjust the quality of the good to the recommender policy the platform is employing. And the platform, in turn, should take the strategic nature of alternatives into account once developing the policy. To my best knowledge, there are no papers that study the optimal recommender policy with strategic alternative, so this work may be considered as the first small step in this direction.

	I build a model that embraces three main counterparties: the platform, the firm and users. Users are choosing between the good of \textit{ex-ante} unknown quality provided by the firm and a safe outside option. The firm can exert costly private effort to increase the quality of the service and the platform design the optimal recommender policy taking this into account. I demonstrate that the optimal policy is different once the firm becomes strategic. Also, I derive the optimal recommender policy in the class of simple "1-round" policies. The rest of the paper is organized as follows: the next section presents the literature review, section 3 describes the model, section 4 contains the main results, and last three sections embrace a brief discussion, possible extensions and directions for the future research, and a short conclusion.
	\section{Literature Review}
	This work is related to the strand of literature on recommender systems. One of the first papers which studied the optimal design of recommender system is the work by Kremer et al \cite{kremer2014}. The authors build a simple model with two alternatives which generate deterministic but \textit{ex-ante} unknown payoff. One of the alternatives is \textit{a priori} better, in the sense that the expected reward it generates is higher. The finite number of short-living consumers arrive sequentially and each chooses between either of two alternatives. The platform accumulates review and commits to the recommendation policy it uses to make recommendations to the newcomers. The papers, firstly, neatly in simple terms demonstrates why the full revelation policy may be suboptimal. Secondly, the optimal recommender policy is derived which has a "threshold" nature. That is, \textit{a priori} better alternative is recommended if its quality turns out to be higher than the certain threshold. Otherwise the other alternative is recommended, and then once both alternatives were tried the best one is recommended until the end. The setup of this model is very similar to my, however, it has a clear distinction. In my model the quality of the alternative is endogenous to the process of recommendation. This suggests that the threshold-like type of policies cannot be optimal in the presence of strategic alternatives because once the firm knows that it will be recommended forever, it does not have any incentives to maintain the high quality of the service it provides.

	Another relevant paper is that of Che and Horner \cite{che2015}. In the paper a continuous flow of customers arrives, and each should choose between the good or an outside option. The quality of the good is exogenously predetermined but unknown. Customers who decide to consume the good send noisy signal about its quality to the platform. The signal is noisy in the sense that even if the quality of the good is good the information about it reaches to the platform only with some probability (different from 1). The platform updates its beliefs about the quality and tailors recommender policy contingent on the beliefs. Authors demonstrate that the full revelation policy is suboptimal and find the optimal policy which again has a threshold nature. Namely, once the platform receives the signal that the quality is good, it will recommend the good until the end. Otherwise, if no news arrive it is optimal to experiment with the risky alternative (recommend it) until the beliefs that its quality is good fall below the certain threshold. After that the optimal policy never recommends this alternative.

	Finally, Papanstasiou et al \cite{papanastasiou2017} extended ideas developed in \cite{kremer2014} and \cite{che2015} and build a dynamic, stochastic model. In their model, agents also arrive sequentially to the platform and should choose between two alternatives that generate stochastically exogenous payoff. The platform in the paper designs a recommendation policy. Although the optimal policy turns out to have a complicated form, authors demonstrate that in this case the full revelation policy is not the best one, and the optimal policy involves information obfuscation.

	The common feature of all aforementioned models is some form of exogeneity of the quality of the alternative. In the next section I present a model with the strategic firm that can adjust the quality of the good to the learning process and recommendation policy.

	\section{Model}
	In the model one firm is operating on the platform and provides customers with the service of ex-ante unknown quality. Unit mass of customers arrives sequentially, each customer decides between the risky good provided by the firm and the safe outside option (customers differ in their outside option). Customers base their decision on the recommendation that is sent by the recommender system. Those who consume the good report the quality to the platform and leave forever. The platform accumulates reviews and commits to the recommender policy that can be a function of past reviews.

	Three subsequent subsections in details describe all parties of the model.
	\subsection{Firm}
	The quality of the good $\mathcal{Q}_t$ provided by the firm at period $t$ can be either good or bad, $\mathcal{Q}_t \in \{g, b\}$. I will refer to the quality of the good as "the state of the firm". Following this notation, the firm can be either in good state or in bad. The firm can exert costly effort which can increase the probability to remain in the good state (if the current state is good) or to move to the good state (if the current state is bad). Without efforts with probability 1 firms either remains in the bad state (if the current state is bad) or moves to the bad state (if the current state is good). At each state the firm chooses the level of effort $e\in\{0, 1\}$ which will be exerted by the firm before the customers arrive. For example, if the firms chooses effort only based on its current state, then $e = \{e^b, e^g\}$. Using the notation from the theory of Markov chains, the transitory matrix in this case looks as follows $$P = \begin{pmatrix}
	1 - pe^{b} & pe^{b}\\
	1-pe^{g} & pe^{g}
	\end{pmatrix}$$
	here the first row refers to the bad state and the second to the good state, $p$ is the exogenous parameter which reflects the probability that investment will be successful. For example, the element $(1, 2)$ of the matrix represents the probability that the firm will move to the state $g$ if its current state is $b$.

	If at time $t$ the share of the customers $s_t \in [0, 1]$ decides to consume the good then the firm gets $$\pi_t = s_t - ce$$ where $c$ is the cost of effort and $e \in \{0, 1\}$. The firm's objective is to choose an investing strategy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \pi_t \right]$$ subject to the recommender policy implemented by the platform.
	\subsection{Customers}

	The unit mass of heterogeneous customers arrives sequentially each period. Each customer has the outside option $r_i \in [0, 1]$. Upon consuming the good customers derive the utility $$v_t = \begin{cases}
	1, \text{ if }\mathcal{Q}_t = g\\
	q, \text{ if }\mathcal{Q}_t = b
	\end{cases}$$
	where $q \in (0, 1)$. The total utility of the customer who has an outside option $r_i$ and arrives at time $t$ is determined as follows $$u_{it} = \begin{cases}
	v_t &,\text{if he chose the good}\\
	r_{i} &,\text{if he chose an outside option}
	\end{cases}$$
	Choosing between outside option and the good, customers take into account a private signal received from the platform. This signal has a form of recommendation: the platform either recommends the customer trying the good or not. 
	\subsection{Platform}
	At the beginning of the game the platform commits to the recommendation policy, that can be based on past reviews it collected. Then, each period the platform recommends the good to the share of customers $s_t$. Where $s_t$ is the mapping from the set of possible histories at time $t$ to $[0, 1]$. If at time $t$ the platform recommends the good to the share of customers $s_t$ and everybody follows the recommendation then the platform derives a utility $$\mathcal{U}_t = v_ts_t - \frac{s_t^2}{2}$$ which is basically an aggregate surplus obtained by customers who chose a good. The platform designs the recommendation policy that would maximize $$\expect_0\left[\sum_{t=0}^{\infty} \delta^t \mathcal{U}_t\right]$$ subject to the firm's strategy $\sigma(s_t(h^t))$ which, in turn, is induced by the policy.
	\subsection{Timing}
	Timing of each period of the game is the following. At the beginning of the period the firm decides on the level of effort it will exert. Then this level is exerted and, the firm privately learns its new state. After that the platform makes a recommendation, customers, who followed it, report reviews and the round ends. 
	\begin{center}
	\begin{figure}[H]
		\begin{tikzpicture}
	\draw[ultra thick,->] (0,0) -- (11,0);
	\draw[thick] (1,-0.2) -- (1,0.2);
	\node[label=below:firm learns] at (1,-0.2) {};
	\node[label=below:its state] at (1,-0.5) {};
	\draw[thick] (3,-0.2) -- (3,0.2);
	\node[label=above:firm makes] at (3,0.5) {};
	\node[label=above:private investment] at (3,0.1) {};
	\draw[thick] (5,-0.2) -- (5,0.2);
	\node[label=below:customers] at (5,-0.2) {};
	\node[label=below:arrive] at (5,-0.5) {};
	\draw[thick] (7,-0.2) -- (7,0.2);
	\node[label=above:a recommendation] at (7,0.2) {};
	\node[label=above:platform makes] at (7,0.5) {};
	\draw[thick] (9,-0.2) -- (9,0.2);
	\node[label=below:platform receives] at (9,-0.2) {};
	\node[label=below:reviews] at (9,-0.5) {};
		\end{tikzpicture}
	\caption{Timing}\label{fig1}
	\end{figure}
\end{center}
 	\subsection{Assumptions}
 	Before starting the analysis I would like to explicitly state assumptions the model is based on, and to provide some justifications of these assumptions.
 	\begin{assumption}
 		The platform knows customers' outside. options
 	\end{assumption}
 	This assumption at first glance may seem implausible, however, nowadays platforms have an enormous amount of personal data which can enable them to infer rather precisely about the potential outside option of each customer.

 	\begin{assumption}
 		If the good is not recommended to the customer, then the customer does not consume it. While if the good is recommended the customer may or may not choose it.
 	\end{assumption}
 	This assumption implies that without the platform the customer is unable to find the good. Think about \textit{TripAdvisor}, if it does not show you a restaurant, you will not know about the existence of this restaurant. At the same time, if \textit{TripAdvisor} shows you an option it is up to you whether to go there or not.
 	\begin{assumption}
 		The utility customers obtain in the bad state, $q > 0$ and hence it is always profitable for the platform to recommend the good to at least a small share of customers.
 	\end{assumption}
 This assumption ensures that the platform can get the information about the state of the firm each period.
 \begin{assumption}
 	$p \ge 0.5$
 \end{assumption}
This assumption, on one hand, is pure technical, and simplifies an analysis of optimal policy, but on the other hand it seems reasonable that once the firm invests, the probability that investment will lead to an increase in quality is at least 50\%.
\section{Analysis}
Although the recommender policy the platform employs can be a complicated function of all previous states, I start the analysis from the simplest policy. Namely, I firstly consider the policy that takes into account only the previous realization of the quality of the good.
\subsection{One-round recommendation policy}
Suppose that the platform employs the recommender policy that is based only on the previous period realization of the quality of the good. In this case the policy consists of two shares $\{s^0, s^1\}$ such that $$s_t = \begin{cases}
s^0, \text{ if }\mathcal{Q}_{t-1} = b\\
s^1, \text{ if }\mathcal{Q}_{t-1} = g
\end{cases}$$
The platform should choose $s^0, s^1$ to maximize its objective taking into account incentives of the firm. But firstly as a benchmark consider the first-best case, when the firm always invest. In this case the recommender policy is characterized by the following lemma

\begin{lemma}(First-best)
	If the firm always invests in quality then the optimal recommender policy implies $$s^1_{FB} = s^0_{FB} = p + (1-p)q$$
\end{lemma}
\begin{proof}
	The proof is straightforward, if the firm invests then the expected quality of the service at each period is $p+(1-p)q$. The platform is maximizing $$\underset{s}{\max}\ (p+(1-p)q)s - \frac{s^2}{2}$$ and the solution is given by $$s = p + (1-p)q$$
\end{proof}
Clearly, if the platform implements this policy, the firm does not have incentives to invest in quality. 
\begin{corollary}
	If the firm is strategic then the first-best policy is not implementable i.e. if the platform employs the first-best recommender policy, the firm never invests.
\end{corollary}

The lemma and the corollary above suggest that the optimal policy should induce the firm to invest in quality. The results below give a characterization of the optimal recommender policy.
\begin{lemma}
	Given the recommendation policy $\{s^0, s^1\}$, it is optimal for the firm to invest at the bad state if and only if it is optimal to invest at the good state.
\end{lemma} 
\begin{proof}
	Let $V_{ij}$ be the expected discounted value of the firm which is at state $i$ and in the previous period was at state $j$. Then this values are the solutions to the following system of equations:
	\begin{align}\label{eq1}
	\begin{cases}
	V_{bb} = s^0 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
	V_{bg} = s^1 + \max\{\delta V_{bb}, \delta p V_{gb} + \delta (1-p)V_{bb} - c\}\\
	V_{gb} = s^0 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}\\
	V_{gg} = s^1 + \max\{\delta V_{bg}, \delta p V_{gg} + \delta (1-p)V_{bg} - c\}
	\end{cases}
	\end{align}
	In each $\max$ the first component represents the case when the firm does not invest, while the second is the value the firm obtains if invests.
